---
title: "Introduction to R"
author: "Shivani Mehta"
date: "1/27/2020"
output: html_document
---

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Calculate total gains for poker and roulette
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)
```



```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Define a new variable based on a selection Tuesday to Friday
roulette_selection_vector <- roulette_vector[c(2:5)]

```


```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
```


```{r}
# Which days did you make money on roulette?
selection_vector <- roulette_vector >0

# Select from roulette_vector these days
roulette_winning_days <- roulette_vector[selection_vector]

```

#Use c(new_hope, empire_strikes, return_jedi) to combine the three vectors into one vector. Call this vector box_office.
#Construct a matrix with 3 rows, where each row represents a movie. Use the matrix() function to do this. 
#The first argument is the vector box_office, containing all box office figures. Next, you'll have to specify nrow = 3 and byrow = TRUE. Name the resulting matrix star_wars_matrix.

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow =TRUE)
```

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region
# Name the rows with titles
rownames(star_wars_matrix) <- titles
star_wars_matrix
# Print out star_wars_matrix

# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), c("US", "non-US")))
# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)
# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)

#Adding a row
box_office2 <- c(411.998, 304.4, 276.475, 247.900, 309.306, 165.8)
star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE,
                           dimnames = list(c("The Phantom Menace", "The Phantom Menace", "Revenge of the Sith"), c("US", "non-US")))
# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
```

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix
# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
# Average non-US revenue
mean(non_us_all)
# Select the non-US revenue for first two movies
non_us_some <- non_us_all[1:2]
# Average non-US revenue for first two movies
mean(non_us_some)
# Estimate the visitors
visitors <- all_wars_matrix/5
# US visitors
us_visitors <- visitors[,1]
# Average number of US visitors
mean(us_visitors)

```
#FACTORS
#The term factor refers to a statistical data type used to store categorical variables. 
#The difference between a categorical variable and a continuous variable is that a categorical variable can belong to a limited number of categories. 
#A continuous variable, on the other hand, can correspond to an infinite number of values.

```{r}
#Nominal
# Sex vector
sex_vector <- c("Male", "Female", "Female", "Male", "Male")
# Convert sex_vector to a factor
factor_sex_vector <- factor(sex_vector)
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <- c("Female", "Male")

factor_survey_vector
```

```{r}
# Ordinal 
#Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector


# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")

# Convert speed_vector to ordered factor vector
factor_speed_vector <- factor(speed_vector, ordered=TRUE, levels=c("slow", "medium", "fast"))

# Print factor_speed_vector
factor_speed_vector
summary(factor_speed_vector)
```
Going back to our survey, you would like to know how many "Male" responses you have in your study, and how many "Female" responses. The summary() function gives you the answer to this question.
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector
# Generate summary for survey_vector
summary(survey_vector)
# Generate summary for factor_survey_vector
summary(factor_survey_vector)

# Male
male <- factor_survey_vector[1]
# Female
female <- factor_survey_vector[2]
# Battle of the sexes: Male 'larger' than female?
male > female
```
```{r}
# Factor value for second data analyst
da2 <-factor_speed_vector[2]

# Factor value for fifth data analyst
da5 <-factor_speed_vector[5]

# Is data analyst 2 faster than data analyst 5?
da2 > da5

```
#Creating a Data Frame
```{r}
# Definition of vectors
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <-data.frame(name, type, diameter, rotation, rings)
planets_df
```

```{r}
# Print out diameter of Mercury (row 1, column 3)
planets_df[1,3]

# Print out data for Mars (entire fourth row)
planets_df[4,]

# Select first 5 values of diameter column
planets_df[1:5,"diameter"]

# Select the rings variable from planets_df
rings_vector <- planets_df$rings
  
# Print out rings_vector
rings_vector

# Adapt the code to select all columns for planets with rings
planets_df[rings_vector,]

# Select planets with diameter < 1
subset(planets_df, subset = diameter < 1)
```

```{r}
# Use order() to create positions
positions <- order(planets_df$diameter)

# Use positions to sort planets_df
planets_df[positions, ]
```
#List

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
```

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Adapt list() call to give the components names
my_list <- list(my_vector, my_matrix, my_df)
names(my_list) <- c("vec","mat","df")
my_list

# Print out my_list
```

```{r}
```

```{r}
```

```{r}
```