---
title: "Introduction to R"
author: "Shivani Mehta"
date: "1/27/2020"
output: html_document
---

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Calculate total gains for poker and roulette
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)
```



```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Define a new variable based on a selection Tuesday to Friday
roulette_selection_vector <- roulette_vector[c(2:5)]

```


```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
```


```{r}
# Which days did you make money on roulette?
selection_vector <- roulette_vector >0

# Select from roulette_vector these days
roulette_winning_days <- roulette_vector[selection_vector]

```

#Use c(new_hope, empire_strikes, return_jedi) to combine the three vectors into one vector. Call this vector box_office.
#Construct a matrix with 3 rows, where each row represents a movie. Use the matrix() function to do this. 
#The first argument is the vector box_office, containing all box office figures. Next, you'll have to specify nrow = 3 and byrow = TRUE. Name the resulting matrix star_wars_matrix.

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow =TRUE)
```

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region
# Name the rows with titles
rownames(star_wars_matrix) <- titles
star_wars_matrix
# Print out star_wars_matrix

# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), c("US", "non-US")))
# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)
# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)

#Adding a row
box_office2 <- c(411.998, 304.4, 276.475, 247.900, 309.306, 165.8)
star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE,
                           dimnames = list(c("The Phantom Menace", "The Phantom Menace", "Revenge of the Sith"), c("US", "non-US")))
# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
```

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix
# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
# Average non-US revenue
mean(non_us_all)
# Select the non-US revenue for first two movies
non_us_some <- non_us_all[1:2]
# Average non-US revenue for first two movies
mean(non_us_some)
# Estimate the visitors
visitors <- all_wars_matrix/5
# US visitors
us_visitors <- visitors[,1]
# Average number of US visitors
mean(us_visitors)

```
#FACTORS
#The term factor refers to a statistical data type used to store categorical variables. 
#The difference between a categorical variable and a continuous variable is that a categorical variable can belong to a limited number of categories. 
#A continuous variable, on the other hand, can correspond to an infinite number of values.

```{r}
#Nominal
# Sex vector
sex_vector <- c("Male", "Female", "Female", "Male", "Male")
# Convert sex_vector to a factor
factor_sex_vector <- factor(sex_vector)
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <- c("Female", "Male")

factor_survey_vector
```

```{r}
# Ordinal 
#Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector


# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")

# Convert speed_vector to ordered factor vector
factor_speed_vector <- factor(speed_vector, ordered=TRUE, levels=c("slow", "medium", "fast"))

# Print factor_speed_vector
factor_speed_vector
summary(factor_speed_vector)
```
Going back to our survey, you would like to know how many "Male" responses you have in your study, and how many "Female" responses. The summary() function gives you the answer to this question.
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector
# Generate summary for survey_vector
summary(survey_vector)
# Generate summary for factor_survey_vector
summary(factor_survey_vector)

# Male
male <- factor_survey_vector[1]
# Female
female <- factor_survey_vector[2]
# Battle of the sexes: Male 'larger' than female?
male > female
```
```{r}
# Factor value for second data analyst
da2 <-factor_speed_vector[2]

# Factor value for fifth data analyst
da5 <-factor_speed_vector[5]

# Is data analyst 2 faster than data analyst 5?
da2 > da5

```
#Creating a Data Frame
```{r}
# Definition of vectors
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <-data.frame(name, type, diameter, rotation, rings)
planets_df
```

```{r}
# Print out diameter of Mercury (row 1, column 3)
planets_df[1,3]

# Print out data for Mars (entire fourth row)
planets_df[4,]

# Select first 5 values of diameter column
planets_df[1:5,"diameter"]

# Select the rings variable from planets_df
rings_vector <- planets_df$rings
  
# Print out rings_vector
rings_vector

# Adapt the code to select all columns for planets with rings
planets_df[rings_vector,]

# Select planets with diameter < 1
subset(planets_df, subset = diameter < 1)
```

```{r}
# Use order() to create positions
positions <- order(planets_df$diameter)

# Use positions to sort planets_df
planets_df[positions, ]
```
#List

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
```

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Adapt list() call to give the components names
my_list <- list(my_vector, my_matrix, my_df)
names(my_list) <- c("vec","mat","df")
my_list

# Print out my_list
```


#Comparing Vectors


```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin > 15

# Quiet days
linkedin <= 5

# LinkedIn more popular than Facebook
linkedin > facebook

linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
15 < last & last <= 20
```

#Compare Matrices

```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)
views
# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <=14
```

#Conditionals

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
    print("You are popular!")
}


# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}else{
  print("Unknown medium") 
}

```

#While

```{r}
# Initialize the speed variable
speed <- 64

# Extend/adapt the while loop
while (speed > 30) {
  print(paste("Your speed is",speed))
  if (speed > 48 ) {
    print ("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```

#Break

```{r}
# Initialize i as 1 
i <- 1

# Code the while loop
while (i <= 10) {
  print(3 * i)
  if ( (3*i)%%8 ==0 ) {
    break
  }
  i <- i + 1
}
```

#For

```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# loop version 1
for (p in linkedin) {
  print(p)
}

# loop version 2
for (i in 1:length(linkedin)) {
  print(linkedin[i])
}
```

#Functions

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
# Calculate average number of views
avg_li = mean(linkedin)
avg_fb = mean(facebook)
# Inspect avg_li and avg_fb
avg_li
avg_fb

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum =mean(linkedin+facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed = mean(linkedin+facebook, trim = 0.2)

# Inspect both new variables
avg_sum
avg_sum_trimmed
```

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Basic average of linkedin
mean(linkedin)

# Advanced average of linkedin
mean(linkedin, trim = 0, na.rm = TRUE)

```

```{r}
# Create a function pow_two()
pow_two <- function(arg1){
  return (arg1*arg1)
}

# Use the function
pow_two(12)

# Create a function sum_abs()
sum_abs <- function(arg1, arg2){
  return (abs(arg1)+abs(arg2))
}


# Use the function
sum_abs(-2,3)
```


```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
# Define the interpret function
interpret <- function(num_views) {
  if (num_views > 15) {
      print( "You're popular!")
      return (num_views)

  } else {
    print("Try to be more visible!")
    return (0)
  }
}

# Call the interpret function twice
interpret(linkedin[1])
interpret(facebook[2])
```

#lapply()
lapply takes a vector or list X, and applies the function FUN to each of its members

```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  x[1]
}

# Apply select_first() over split_low: names
names<-lapply(split_low, select_first)
# Write function select_second()
select_second <- function(x) {
  x[2]
}
# Apply select_second() over split_low: years
years<-lapply(split_low, select_second)

# Transform: use anonymous function inside lapply

lapply(split_low, function(x) { x[1] })

# Transform: use anonymous function inside lapply

lapply(split_low, function(x) { x[2] })


# Definition of split_low
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Generic select function
select_el <- function(x, index) {
  x[index]
}

# Use lapply() twice on split_low: names and years
names<-lapply(split_low, select_el,index=1)
years<-lapply(split_low, select_el,index=2)
```


#sapply() Simplifies the lapply
```{r}

temperature <- c(3,7,9,6,-1,6,9,12,13,5,4,8,3,-1,-3,1,4,7,2,-2,5,7,9,4,2,-3,5,8,9,4,3,6,9,4,1)
 temp <- matrix(temperature, nrow = 7, byrow =TRUE)
# temp is already prepared for you in the workspace

# Definition of below_zero()
below_zero <- function(x) {
  return(x[x < 0])
}

# Apply below_zero over temp using sapply(): freezing_s
freezing_s <- sapply(temp, below_zero)

# Apply below_zero over temp using lapply(): freezing_l
freezing_l <- lapply(temp, below_zero)

# Are freezing_s and freezing_l identical?
identical(freezing_s, freezing_l)


# Definition of print_info()
print_info <- function(x) {
  cat("The average temperature is", mean(x), "\n")
}

# Apply print_info() over temp using sapply()
sapply(temp,print_info)

# Apply print_info() over temp using lapply()
lapply(temp,print_info)


# temp is already defined in the workspace

# Convert to vapply() expression
vapply(temp, max, FUN.VALUE=numeric(1))

# Convert to vapply() expression
vapply(temp, function(x, y) { mean(x) > y }, y = 5, FUN.VALUE=logical(1))
```

R features a bunch of functions to juggle around with data structures::

seq(): Generate sequences, by specifying the from, to, and by arguments.
rep(): Replicate elements of vectors and lists.
sort(): Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.
rev(): Reverse the elements in a data structures for which reversal is defined.
str(): Display the structure of any R object.
append(): Merge vectors or lists.
is.*(): Check for the class of an R object.
as.*(): Convert an R object from one class to another.
unlist(): Flatten (possibly embedded) lists to produce a vector.

```{r}
```

grepl(), which returns TRUE when a pattern is found in the corresponding character string.
grep(), which returns a vector of indices of the character strings that contains the pattern.

```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for "edu"
grepl(pattern="edu", x=emails)

# Use grep() to match for "edu", save result to hits
hits<-grep(pattern="edu", x=emails)

# Subset emails using hits
emails[hits]


# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for .edu addresses more robustly
grepl(pattern="@.*\\.edu$", x=emails)

# Use grep() to match for .edu addresses more robustly, save result to hits
hits = grep(pattern="@.*\\.edu$", x=emails)

# Subset emails using hits
emails[hits]
```

#sub() only replaces the first match, whereas gsub() replaces all matches.
```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "global@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use sub() to convert the email domains to datacamp.edu
sub(pattern="@.*\\.edu$", replacement = "@datacamp.edu", x= emails) 
```

Before you rack your brains over the regular expression in this exercise, have a look at the new things that will be used:

.*: A usual suspect! It can be read as "any character that is matched zero or more times".
\\s: Match a space. The "s" is normally a character, escaping it (\\) makes it a metacharacter.
[0-9]+: Match the numbers 0 to 9, at least once (+).
([0-9]+): The parentheses are used to make parts of the matching string available to define the replacement. The \\1 in the replacement argument of sub() gets set to the string that is captured by the regular expression [0-9]+.

```{r}
```


Time and Date

%Y: 4-digit year (1982)
%y: 2-digit year (82)
%m: 2-digit month (01)
%d: 2-digit day of the month (13)
%A: weekday (Wednesday)
%a: abbreviated weekday (Wed)
%B: month (January)
%b: abbreviated month (Jan)
%H: hours as a decimal number (00-23)
%I: hours as a decimal number (01-12)
%M: minutes as a decimal number
%S: seconds as a decimal number
%T: shorthand notation for the typical format %H:%M:%S

```{r}
# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2 <- as.Date(str2)
date3 <- as.Date(str3, format = "%d/%B/%Y")

# Convert dates to formatted strings
format(date1, "%A")
format(date2, "%d")
format(date3, "%b %Y")


# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2)

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```
