---
title: "Assignment1"
author: "Group 5 - 
 Sourabh Bodkhe,
 Manali Thorat,
 Shivani Mehta,
 Jatan Rathod,
 Anvay Thakare"
date: "2/7/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# *Assignment 1 *


### Load packages and check loading  
```{r loadPackages, warning=FALSE, message=FALSE, results='hide'}
#Importing the Utilities.csv in to a data table
if(!require("data.table")) install.packages("data.table")

utilities.dT<-fread("Utilities.csv")

```


1.Compute the minimum, maximum, mean, median, and standard deviation for
each of the numeric variables using data.table package. Which variable(s) has the largest variability? Explain your answer.

```{r}
#Mean
utility_mean<-utilities.dT[, lapply(.SD, mean),.SDcols = c("Fixed_charge","RoR","Cost","Load_factor",	"Demand_growth","Sales","Nuclear",	"Fuel_Cost")]
#Median
utility_median<-utilities.dT[, lapply(.SD, median), .SDcols = c("Fixed_charge","RoR","Cost","Load_factor",	"Demand_growth","Sales","Nuclear",	"Fuel_Cost")]
#Standard Deviation
utility_sd<-utilities.dT[, lapply(.SD, sd), .SDcols = c("Fixed_charge","RoR","Cost","Load_factor",	"Demand_growth","Sales","Nuclear",	"Fuel_Cost")]
#Minimum
utility_min<-utilities.dT[, lapply(.SD, min), .SDcols = c("Fixed_charge","RoR","Cost","Load_factor",	"Demand_growth","Sales","Nuclear",	"Fuel_Cost")]
#Maximum
utility_max<-utilities.dT[, lapply(.SD, max), .SDcols = c("Fixed_charge","RoR","Cost","Load_factor",	"Demand_growth","Sales","Nuclear",	"Fuel_Cost")]
  utilities_stats_matrix <- matrix(c(utility_mean, utility_median, utility_sd, utility_min, utility_max), nrow = 5, ncol = 8, byrow = TRUE)
  colnames(utilities_stats_matrix) <- c("Fixed_Charge", "RoR", "Cost", "Load_factor", "Demand_growth", "Sales", "Nuclear", "Fuel_Cost")
  rownames(utilities_stats_matrix)<-(c("Mean", "Median", "StdDev","Minimum", "Maximum"))
utilities_stats_matrix
```
Standardizing to find actual variance in the data

```{r}
coefficient.variation<-utility_sd/utility_mean
coefficient.variation
```
Interpretation:
From the output above, we see Nuclear have the highest coefficient of variation followed by Demand_growth, Fuel_Cost, Sales,Cost, RoR, Fixed_Charge and Load_Factor.Since high coefficient of variation suggest higher variability, we can conclude that Nuclear has the largest variability.


2. Create boxplots for each of the numeric variables. Are there any extreme values for any of the variables? Which ones? Explain your answer.

```{r}
if(!require("ggplot2")) install.packages("ggplot2")
ggplot(stack(utilities.dT[,2]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Fixed Charge variable ")

ggplot(stack(utilities.dT[,3]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for RoR variable ")

ggplot(stack(utilities.dT[,4]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Cost variable ")

ggplot(stack(utilities.dT[,5]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Load_Factor variable ")

ggplot(stack(utilities.dT[,6]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Demand_growth variable ")

ggplot(stack(utilities.dT[,7]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Sales variable ")

ggplot(stack(utilities.dT[,8]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Nuclear variable ")

ggplot(stack(utilities.dT[,9]), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title= "  Boxplot for Fuest_Cost variable ")

```

Interpretation:
Yes, Sales and Fixed Charge variables have extreme values, as we can see from the points lying outside the ends of the whiskers which is 1.5*InterQuartile Range also called outliers.
Values of Outliers for Sales: 17441, 15991
Values of Outliers Fixed Chargs: 1.49, 1.43, 0.75 and 0.76

3.Create a heatmap for the numeric variables. Discuss any interesting trend you see in this chart.

```{r Heatmaps, warning=FALSE, message=FALSE}
library(reshape)
cor.mat <- round(cor(utilities.dT[,!c("Company")]),2) #rounded correlation matrix 
melted.cor.mat <- melt(cor.mat) 
ggplot(melted.cor.mat, aes(x = X1, y = X2, fill = value)) + 
  scale_fill_gradient(low="wheat", high="orangered") +
  geom_tile() + 
  geom_text(aes(x = X1, y = X2, label = value)) +
  ggtitle("Which Variables Are Highly Correlated?")
```
Interpretation:
1. The darker color in the heat map shows higher correlation between the two variables, here RoR has a positive correlation with Fixed_charge of 0.64 which is closer to +1.
2.Similary, Sales and Fuel_Cost have a negative correlation of -0.56 which is closer to -1.
3. Fuel_Cost and cost have coefficient of correlation is 0.01 ~ 0, hence there is no liner relationship between the two.


4. Run principal component analysis using unscaled numeric variables in the
dataset. How do you interpret the results from this model?


```{r}
#PCA on all components
pca <- prcomp(na.omit(utilities.dT[,!c("Company")]))
summary(pca)
#Performing rotational matrix
pca$rotation

```
Interpretation:
Without standardizing the data set principal component 1 (PC1) alone accounts towards 99.98% of the variance. But in order to give equal emphasis on each variable we should scale the data.


5. Next, run principal component model after scaling the numeric variables. Did the results/interpretations change? How so? Explain your answers.

```{r}
pca.cor <- prcomp(na.omit(utilities.dT[,!c("Company")]), scale. = T)# Normalizing the data
summary(pca.cor)
#Performing rotational matrix
pca.cor$rotation

```
Interpretation:
Without standardizing the result only 1 principal component was counted towards whole variance but after normalizing the data we can see that each principal component accounts for some variance. Even if we consider first 6 principal components, we get total variance as 95.11.
